---
title: "630_hw3_amber"
author: "Ruohan wang"
date: "3/6/2022"
output:
  pdf_document: default
  html_document: default
---

```{r}
#630 hw3
#1 a
setwd("~/Downloads")
df1 = read.csv("pancreatic.csv")
df1$entry_date = as.Date(df1$entry_date,"%m/%d/%y")
df1$progression_date= as.Date(df1$progression_date,"%m/%d/%y")
df1$death_date= as.Date(df1$death_date,"%m/%d/%y")
#df1
#estimate the median PFS in months using R
PFS=c()
for(i in 1:41){
  if(is.na(df1$progression_date[i])==TRUE){
    PFS[i] = df1$death_date[i] - df1$entry_date[i]
  }else if(df1$progression_date[i] > df1$death_date[i]){
    PFS[i] =  df1$death_date[i] - df1$entry_date[i]
  }else if(df1$progression_date[i] < df1$death_date[i]){
    PFS[i] = df1$progression_date[i] - df1$entry_date[i]
  }
} 
PFS_month = PFS/30
cbind(df1,PFS_month)
month = median(PFS)/30
month
```

Therefore the median PFS in months is 3.966667 months.

```{r}
#1b
library(survival)
library(survminer)

df1$d3 <- 1
fit <- survfit(Surv(PFS_month, df1$d3) ~ 1)
breaks <- seq(6, 18, 6)
hzd <- c(0.2, 0.15, 0.15, 0.18)
t <- seq(0, 30, 0.01)
# locate each element of t in breaks
idx <- findInterval(t, breaks, left.open = T) + 1
H <- cumsum(hzd[idx]*0.01)
S <- exp(-H)
par(mfrow = c(1,1), mar = c(4, 4, 1, 1))
plot(fit, conf.int = FALSE, lwd = 2, lty = 1,
     mark.time = T, las = 1, xlab = "Months",  ylab = "Survival")
lines(t, S, col = 3, lwd = 2)
abline(h = 0.5, col = "grey", lty = 2)
```
```{r}
#1 c
lamda = c(0.2, 0.15, 0.18)
survdiff(Surv(PFS_month) ~ offset(exp(-(lamda[1]*PFS_month + (lamda[2]-lamda[1])*(PFS_month-6)*I(PFS_month > 6
) + (lamda[3] - lamda[2])*(PFS_month - 18)*I(PFS_month>18)))), data = df1)
```
Since the p value is 0.571 and is greater than 0.05, so we fail to reject the null hypothesis.

Therefore, there is no different from the survival curve of SOC.


```{r}
#3 a
library(KMsurv)
data("tongue")
source("FH test.R")
source("renyi.R")

reg <- survdiff(Surv(time, delta) ~ type, rho = 0 , data = tongue) 

p= 1- pchisq(q = reg$chisq, df = 1)
p
```
The p value is 0.09 and is greater than 0,05, so this is large.

Fail to reject H0, so survival rates of patients with cancer of the tongue for patients with aneuploid is the same as the patients with diploid tumors.

```{r}
#3 b
reg.1 <- survdiff(Surv(time, delta) ~ type, rho = 1, data = tongue)
p.1=1- pchisq(q = reg.1$chisq, df = 1)
p.1
```
The p value is 0.069 and is greater than 0,05, so this is large.

Fail to reject H0, so there is no evidence of detecting differences in survival rates be-
tween the two types of cancers which occur soon after the diagnosis of the cancer.

```{r}
#3 c

tongue$type_1 <- tongue$type - 1
combo.test(t = tongue$time, delta = tongue$delta, x= tongue$type_1)
```
Since the p value is still larger than 0.05, so still fail to reject.

```{r}
#4 a
data("kidtran")
#kidtran
reg4 <- survdiff(Surv(time, delta) ~ race+gender, kidtran)
reg4
fit4 <- survfit(Surv(time, delta) ~ race+gender, kidtran)
ggsurvplot(fit4,  kidtran, title="survival curves",  xlab="Time",
           ylab="Probability of Survival",
           conf.int=F,  axes.offset=F,
           xlim=c(0, 500), ylim=c(0.8, 1),
           break.time.by=50,
           risk.table=T, tables.height=0.3,
           risk.table.title="Patients at Risk",
           risk.table.y.text=F, fontsize=4,
           tables.theme = theme_cleantable(),
           legend=c(0.8, 1),  legend.title="",
           legend.labs=c("white-male", "white-female", "black-male", "black-female"))
```
H0: white males, black males, white females, and black females four groups are the same.
H1: white males, black males, white females, and black females four groups are not the same.

The p value is 0.2 and is greater than 0,05, so this is large.

Fail to reject H0, so white males, black males, white females, and black females four groups are the same.

```{r}
#4 b
reg4.1<-survdiff(Surv(time, delta) ~ strata(gender) + race, data = kidtran)
reg4.1
```
H0: The blacks have a higher mortality rate than whites Adjusting by stratification for the sex of the patient. 
H1: The blacks have no higher mortality rate than whites Adjusting by stratification for the sex of the patient.

The p value is 0.3 and is greater than 0,05, so this is large.

Fail to reject H0, so The blacks have a higher mortality rate than whites Adjusting by stratification for the sex of the patient. 

```{r}
#4 c
# Divide races
male <- kidtran[which(kidtran$gender==1), ] #male
female <- kidtran[which(kidtran$gender==2), ]
# Visual inspection
fit4c <- survfit(Surv(time, delta) ~ race, male)
reg4c <- survdiff(Surv(time, delta) ~ race, male) 
p4c <- 1-pchisq(reg4c$chisq, df=1)
p4c

fit4c2 <- survfit(Surv(time, delta) ~ race, female)
reg4c2 <- survdiff(Surv(time, delta) ~ race, female) 
p4c2 <- 1-pchisq(reg4c2$chisq, df=1)
p4c2

```

H0: Black male/female have the same mortality as whites male/female. 
HA: Blacks male/female have a higher mortality rate than whites male/female.

The p value is 0.02768642. and is less than 0,05, so this is small

Reject H0, so Blacks male/female have a higher mortality rate than whites male/female. 

```{r}
#4 c plot
plot(fit4c, mark.time=T, col=1:2, lwd=2, xlim=c(0, 500), ylim=c(0.80, 1),
     ylab="Probability of Survival", xlab="Time", las=1,
     cex.axis=1.2, cex.lab=1.2, main="Female")
legend("bottomleft", col=1:3, lwd=2, bty="n",
       lty=1, c("White","Black"),
       cex=1)

plot(fit4c2, mark.time=T, col=1:2, lwd=2, xlim=c(0, 500), ylim=c(0.80, 1),
     ylab="Probability of Survival", xlab="Time", las=1,
     cex.axis=1.2, cex.lab=1.2, main="Female")
legend("bottomleft", col=1:3, lwd=2, bty="n",
       lty=1, c("White","Black"),
       cex=1)
```
```{r}
#4 d
for (i in 1:nrow(kidtran)){
  if (kidtran$age[i] <= 33){
    kidtran$newage[i] = 1
  }else if (kidtran$age[i] < 43 ){
    kidtran$newage[i] = 2
  }else if (kidtran$age[i] <= 54){
    kidtran$newage[i] = 3
  }else {
    kidtran$newage[i] = 4
} }
reg4d <- survdiff(Surv(time, delta) ~ newage, rho = 0, data = kidtran)
reg4d
p4d = 1- pchisq(q = reg4d$chisq, df = 3)
p4d
```
since the p value is 4.239487e-11 so this is very significant, then find which pairs are most significant:

```{r}
kid1<- kidtran[which(kidtran$newage==1) , ]
kid2 <- kidtran[which(kidtran$newage==2) , ]
kid3<- kidtran[which(kidtran$newage==3) , ]
kid4 <- kidtran[which(kidtran$newage==4) , ]
kid12<-rbind(kid1, kid2)
reg12 <- survdiff(Surv(time, delta) ~ newage, kid12)
p12 <- 1-pchisq(reg12$chisq, df=1)
p12

kid13<-rbind(kid1, kid3)
reg13 <- survdiff(Surv(time, delta) ~ newage, kid13)
p13 <- 1-pchisq(reg13$chisq, df=1)
p13

kid14<-rbind(kid1, kid4)
reg14 <- survdiff(Surv(time, delta) ~ newage, kid14)
p14 <- 1-pchisq(reg14$chisq, df=1)
p14

kid23<-rbind(kid2, kid3)
reg23 <- survdiff(Surv(time, delta) ~ newage, kid23)
p23 <- 1-pchisq(reg23$chisq, df=1)
p23

kid24<-rbind(kid2, kid4)
reg24 <- survdiff(Surv(time, delta) ~ newage, kid24)
p24 <- 1-pchisq(reg24$chisq, df=1)
p24


kid34<-rbind(kid3, kid4)
reg34 <- survdiff(Surv(time, delta) ~ newage, kid34)
p34 <- 1-pchisq(reg34$chisq, df=1)
p34

p.adjust(c(p12,p13,p14,p23,p24,p34), method = "bonferroni")
```
except the group 34 has pvalue of 0.05906735, all the others have the p value less than 0.05.

So the group 34's age has significant difference in survival.

After Bonferroni method, all the p values are above.




