---
title: "630hw4"
author: "Ruohan wang"
date: "4/16/2022"
output:
  pdf_document: default
  html_document: default
---

```{r}
rm(list=ls(all=TRUE))
library(survival)
library(survminer)
#install.packages("rms")
library(rms)
library(Hmisc)
library(KMsurv)
#' 1
setwd("~/Downloads")
df = read.csv("addicts.csv")
cox.model =coxph(Surv(time, status) ~ clinic+prison+dose, data = df)
summary(cox.model)
dd <- datadist(df)
options(datadist='dd')
for(i in 3:8){
  knot <- cph(Surv(time, status) ~ rcs(dose, i)+strat(clinic)+strat(prison), data=df)
  print(AIC(knot))
} 

#since when knot number =3, the AIC is the smallest.
cox.model2 <- cph(Surv(time, status) ~ rcs(dose, 3) +
clinic+prison, data=df)
anova(cox.model,cox.model2)
```

H0: linearity assumption on dose 

Since the p value is 0.3354 > 0.05, which means fail to reject the h0, the does has linearity assumption.  so choose cox.model.

```{r}
#' 2
##------------------------------------------
## test PH
##------------------------------------------
ph_test <- cox.zph(cox.model, transform = "identity", global = TRUE)
ph_test$table

ph_test <- cox.zph(cox.model, transform = "km", global = TRUE)
ph_test$table

ph_test <- cox.zph(cox.model, transform = "rank", global = TRUE)
ph_test$table

#pdf("test PH.pdf", width=10, height=5)
par(mfrow=c(1, 3))
plot(ph_test)
#dev.off()
```

H0: coxmodel1 obeys PH assumption

From the above plots, we can see the dose fit the PH assumption, first, we can tell from the plots that only the time plot fit the 
pattern of assumption, which displays around the 0. 

Also the p value of clinic is significsnt so clinic is not valid for PH assumption. The prisona and dose obeys PH assumption.


```{r}
#'  3
fit.1 <- survfit(Surv(time/365, status) ~ 1,
                df[df$clinic==1, ]) 
fit.2 <- survfit(Surv(time/365, status) ~ 1,
                df[df$clinic==2, ])
plot(fit.1$time, log(fit.1$cumhaz), type="l",
     xlim = c(0,3), xlab="Time on study (Year)",
     ylab="logH(t)", col=1, lwd=2, las=1)
lines(fit.2$time, log(fit.2$cumhaz), col=2, lwd=2)
legend("bottomright", lty=1, lwd=2, col=1:2,
       c("Clinic 1", "Clinic 2"), bty = "n")
```
```{r}
fit <- survfit(Surv(time/365, status) ~ clinic, data = df)
cox <- coxph(Surv(time/365, status) ~ clinic, data = df)
fit_cox <- survfit(cox, newdata = list(clinic=c(1,2)))
plot(fit, col = 1:2, lwd=2, las=1, xlab="Year")
lines(fit_cox, lwd=2, col=1:2, lty=2)
legend("topright", lty=c(1,1,2,2), col=c(1,2,1,2),
       c("Observed clinic=1", "Observed clinic=2",
         "predicted clinic=1", "predicted clinic=2"),bty="n")
```

H0: clinic obeys PH assumption

For first graph: if the two lines are parallel, then they obeys PH assumption. since they are not, so clinic not obeys PH assumption.

For second graph: since the observed 2 and predicted 2 are not similar, as a result, clinic not obeys PH assumption.

```{r}
#' 4
cox.model3 = coxph(Surv(time, status) ~ tt(clinic)+dose+prison, tt=function(x, t, ...){ifelse(t<=365, 0,x)},
                ties = "breslow", data=df)
summary(cox.model3)

```

H0: bata_ttclinic =0

From above the CI for tt(clinic)x1 is (0.38292 ,1.0416),
CI for tt(clinic)x1 is (0.07542,0.3424). The HR are 0.6316 and
0.1607.  so the clinic 1 has bad performance compared with clinic 2.

```{r}
#' 5
summary(cox.model3)
```

H0: bata_ttclinic =0

Based on the change-point model, does the variable clinic satisfy the PH assumption. since the p for ttclinic is 2.30e-06< 0.05, so reject H0, so the variable clinic does noy satisfy the PH assumption.

```{r}
#' 6
#Define the time-dependent variable as a linear function of time
cox.model6 <- coxph(Surv(time, status) ~ clinic + tt(clinic)++ dose + prison, tt=function(x, t, ...) x*t, ties = "breslow",data=df)

summary(cox.model6)
n =c(100,300,500)
for(i in 1:3){
  t = n[i]
  t2= t^2
  b = cox.model6 $ coef[1]+ t*cox.model6$coef[2]
  v = sqrt(cox.model6 $var[1 , 1] + t2 * cox.model6 $var[2 , 2]
           +2 * t*cox.model6$var[1,2])
  outcome = exp(c(b, b - qnorm(0.975)*v, b + qnorm(0.975)*v))
  cat("t=",t)
  print(outcome)
  
}
```

H0: bata_ttclinic =0

since p value for ttclinic is 0.00143<0.05, so
reject H0, so the variable clinic does not satisfy the PH assumption. then using loop for t =100 300 and 500 days. The HR, 95% CIs for the HRs at these time points are the outcome above.

```{r}
#' 7
#install.packages("survRM2")
library(survRM2)

rmst_reg <- rmst2(time=df$time,status=df$status,arm=ifelse(df$clinic=="2", 1, 0),tau=800,covariates=df[, c(5,6)])
rmst_reg
```

On the first 800 days of follow-up time, the between-group contrast 
is 107.148 =108 days. So clinic 2 can live longer 108days than clinic1. And p = 0.008<0.05, so effect of clinic on survival is significant. 


```{r}
#' 8
reg.exp <- survreg(Surv(time, status) ~ clinic + prison + dose,
                dist = "exponential", data = df)
reg.wei <- survreg(Surv(time, status) ~ clinic + prison + dose,
                dist = "weibull", data = df)

reg.logi <- survreg(Surv(time, status) ~ clinic + prison + dose,
                dist = "loglogistic", data = df)
reg.lono <- survreg(Surv(time, status) ~ clinic + prison + dose,
                dist = "lognormal", data = df)
summary(reg.exp)
summary(reg.wei)
summary(reg.logi)
summary(reg.lono)

AIC(reg.exp)
AIC(reg.wei)
AIC(reg.logi)
AIC(reg.lono)
```

The all the predictor of interest(clinic, prison record, and methadone dose (mg/day)) in your model are above in aummary part.

AIC for reg.wei model is the smallest, so the reg.wei mdoel is the best.

```{r}
#' 9
# summary(reg.exp)
lambda1 <- exp(-reg.exp$linear.predictors)
s <- 1 - pexp(df$time, lambda1)
cs_exp <- -log(s)
fit1 <- survfit(Surv(cs_exp, df$status) ~ 1,
              type = "fleming-harrington")
plot(fit1, fun = "cumhaz", conf.int = F,
     main="Exponential",
     xlab="Cox-Snell Residuals",
     ylab="H(t)",
     xlim=c(0, 2.5), ylim=c(0, 3), las = 1, lwd = 3)
abline(a=0, b=1, col=2, lwd = 3)
# summary(reg.wei)
lambda2 <- exp(-reg.wei$linear.predictors)
alpha2 <- 1/reg.wei$scale
q=-(lambda2 * df$time)**alpha2
s2 <- exp(q)
cs_exp2 <- -log(s2)
fit2 <- survfit(Surv(cs_exp2, df$status) ~ 1,
              type = "fleming-harrington")
plot(fit2, fun = "cumhaz", conf.int = F,
     main="Weibull",
     xlab="Cox-Snell Residuals",
     ylab="H(t)",
     xlim=c(0, 2.5), ylim=c(0, 3), las = 1, lwd = 3)
abline(a=0, b=1, col=2, lwd = 3)
# summary(reg.logi) formula?
lamda3<- exp(-reg.logi$linear.predictors)
s3<-1/(1 + lamda3 * df$time)^(1 / reg.logi$scale)
cs_exp3 <- -log(s3)
fit3 <- survfit(Surv(cs_exp3, df$status) ~ 1, type = "fleming-harrington")
plot(fit3, fun = "cumhaz", conf.int = F,
     main="logistic model",
     xlab="Cox-Snell Residuals",
     ylab="H(t)",
     xlim=c(0, 2.5), ylim=c(0, 2.5), las = 1, lwd = 3)
abline(a = 0, b = 1, col = 2, lwd = 3)
# summary(reg.lono)
alpha4 <- 1 / reg.lono$scale
cs_exp4<- -log(1 - pnorm(alpha4*log(df$time) - reg.lono$linear.predictors*alpha4))
fit4<-survfit(Surv(cs_exp4, df$status) ~ 1,type = "fleming-harrington")
plot(fit4, fun = "cumhaz", conf.int = F,
     main="lognormal",
     xlab="Cox-Snell Residuals",
     ylab="H(t)",
     xlim=c(0, 2.5), ylim=c(0, 3), las = 1, lwd = 3)
abline(a=0, b=1, col=2, lwd = 3)
```


From the plots since the weibull has the most similar patterns of two lines, so weibull model is the best model.

```{r}
#' 10
#reg.wei is the mode we choose
#for clinic 1
predict(reg.wei, newdata = data.frame(clinic=1,prison=1,
                     dose=50), type = "quantile", p = 0.5, se.fit = T)
predict(reg.wei, newdata = data.frame(clinic=2,prison=1,
                     dose=50), type = "quantile", p = 0.5, se.fit = T)

```

Clinic 1's the median survival times is 254.2196 , 

Clinic 1's the median survival times is 516.585.
```{r}
cat("Acceleration factor for c2 is",exp(-reg.wei$coefficients[2]))
```

The median survival times in c1 is 0.4921157 times of c2.


```{r}
summary(reg.wei)
cat("hazard ratio of clinic 2 vs. clinic 1 based on the Weibull model is",exp(-0.70904/exp(-0.31495)))
```

The hazrad rate of c1 and c2 is 0.3785072.
