---
title: "630hw3"
author: "Ruohan wang"
date: "3/25/2022"
output:
  pdf_document: default
  html_document: default
---

```{r}
#1a
library(survival)
library(survminer)
library(KMsurv)
library(xtable)
library(dplyr)
data(hodg)

#summary(hodg)
#dtype is NHL=1 and HOD=2
#gtype is ALLo = 1 Auto =2
for(i in 1:43){
  if(hodg$gtype[i] == 1 & hodg$dtype[i]==1){
    hodg$combine[i] =1
  }else if(hodg$gtype[i] == 1 & hodg$dtype[i]==2){
    hodg$combine[i] =2
  }else if(hodg$gtype[i] == 2 & hodg$dtype[i]==1){
    hodg$combine[i] =3
  }else if(hodg$gtype[i] == 2 & hodg$dtype[i]==2){
    hodg$combine[i] =4
  }
}
time = hodg$time
delta = hodg$delta
combine= as.factor(hodg$combine)

#hodg
cox <- coxph(Surv(time, delta) ~ combine,ties='breslow')
summary(cox)
```


Since the exp(coef) is the hazard ratio. 

The hazard ratio with NHL Allo and HOD Allo is 6.232290, p is 0.006739661. which is very small, so there is a difference exist between two groups. 

The hazard ratio with NHL Allo and NHL Auto is 1.942291, p is 0.239389122 which is not small, so there is no difference exist between two groups.

The hazard ratio with NHL Allo and HOD Auto is 1.942291, p is 0.794058239 which is not small, so there is no difference exist between two groups.

CI for three groups are ( 1.6589, 23.414), (0.6427, 5.870), (0.3677, 3.698). which means there are 95 percent probability that the hazard ratio will between these.
```{r}
#1b
## Score test
round(summary(cox)$sctest, 4)
## Wald test
round(summary(cox)$waldtest, 4)
## likelihood ratio test
round(summary(cox)$logtest,4)
```
H0: beta2=beta3=beta4 = 0 
Ha: at least one of them is not 0

The three tests are different methods to test whether we should choose 
null H0 or alternative Ha. since all three methods have the p-value <0.05
which means significant. we choose to reject H0.

```{r}
#1c
#likelihood test
gtype=as.factor(hodg$gtype)
dtype=as.factor(hodg$dtype)
cox2 = coxph(Surv(time, delta) ~ gtype+dtype+gtype*dtype,ties='breslow')
cox3 =  coxph(Surv(time, delta) ~ gtype+dtype,ties='breslow')
## df=1 for gtype*dtype
lrt <- 2*(cox2$loglik[2]-cox3$loglik[2])
## The p-value of this test suggests that survival is different

p = 1-pchisq(lrt, 1)
p
#wald
var = cox2$var[3,3]
m = t(cox2$coef[3])%*%solve(var)%*%cox2$coef[3]
p2 = 1-pchisq(m,1)
p2
```

since the p value for likelihood is 0.006690241< 0.05. Reject H0, which means there is survival difference between them.

While the wald test shows that the 0.006006927 < 0,05. Reject H0, which means there is some survival difference between them.

```{r}
#1d
summary(cox)
```

since the NHL Allo is the base line. The CI for NHL Auto is (0.6427,5.870).

```{r}
#1e
c1 <- c(1,0,0)
h1 =t(c1)%*%cox$var%*%c1
k1= t(c1%*%cox$coef)%*%solve(h1)%*%(c1%*%cox$coef)
print(1-pchisq(k1, 1))
```

Wald test: 

H0: hazard(HOD_Allo)=hazard(NHL_Allo)
Ha: hazard(HOD_Allo)!=hazard(NHL_Allo)

Function:
For NHL allo: h(t|xi) = h0(t)

For HOD auto: h(t|xi) = h0(t)exp(beta1)

The p-value is 0.006739661< 0.05. Reject H0, which means there is some survival difference between them.

```{r}
#1f
c2 <- c(0,1,-1)
h2 =t(c2)%*%cox$var%*%c2
k2= t(c2%*%cox$coef)%*%solve(h2)%*%(c2%*%cox$coef)
print(1-pchisq(k2, 1))
```
Wald test: 

H0: hazard(HOD Auto)=hazard(NHL Auto )
Ha: hazard(HOD Auto)!=hazard(NHL Auto )

Function:
For HOD Auto: h(t|xi) = h0(t)

For NHL Auto: h(t|xi) = h0(t)exp(beta3)

The p-value is 0.312685>0.05. Fail to reject H0, which means there is no survival difference between them.

```{r}
#1g
c3 <- rbind(c(0,1,0),c(1,0,-1))
k3= t(c3%*%cox$coef)%*%solve(c3%*%cox$var%*%t(c3))%*%(c3%*%cox$coef)
print(1-pchisq(k3, 2))
```

At least one group are different using a two-degree of freedom test of 

H0: h(t|N HLAllo) = h(t|N HLAuto)
H0: h(t|HODAllo) = h(t|HODAuto)

The p-value is 0.01429393<0.05. Reject H0, which means At least one group are different.

```{r}
#2a
library(survival)
library(survminer)
library(KMsurv)
library(xtable)
library(dplyr)
setwd("~/Downloads")
data <- read.csv("leukemia.csv")

ma <- data[data$sex==1, ]
fe <- data[data$sex==0, ]
# 
# time=data$time
# relapse= data$relapse
# trt= data$trt
# logWBC= data$logWBC
reg1 <- coxph(Surv(time, relapse) ~ trt+logWBC, data = ma, ties = "breslow")
summary(reg1)$coef
reg2 <- coxph(Surv(time, relapse) ~ trt+logWBC, data = fe, ties = "breslow")
summary(reg2)$coef
```
Compare the re-gression coefficients from Cox PH models for male and female patients. 

For male group the hazard ratio for trt and logwbc are 0.155821 and 5.149590
treatment group reduce the hazard by 0.155821 which means the trt is a good variable. similar, since HR of logWBC>1 so logWBC is not a good sign, and compared with male, female has stronger change. For female group the hazard ratio for trt and logwbc are 0.7658851 and 3.2223954


```{r}
#2b
coxplot=survfit(coxph(Surv(time,relapse) ~ strata(trt) +logWBC + strata(sex), data = data))
plot(coxplot,col=c(2,3,4,5), xlim=c(0, 35), lty=2, lwd=2, las=2,xlab="time", ylab="survival probability",legend.labs=c("female","male", "female_trt", "male_trt"))
```


From the plot we can tell the effect of treatment differ by gender. 

```{r}
#2c
#data$sex=as.factor(data$sex)
reg3 <- coxph(Surv(time, relapse) ~ trt + logWBC+strata(sex) , data = data, ties = "breslow" )
log1 <- reg3$loglik[2]
log2 <- reg1$loglik[2] + reg2$loglik[2]
diff = 2*(log2 - log1)
p2c<- 1-pchisq(diff,2)
p2c
```

If we use breslow method. The p value is 0.1782153 and >0,05, we fail to reject h0. which means stratification by sex is a good choice by testing whether the
effect of treatment and logWBC differ by sex.
```{r}
#2d
summary(reg3)
```

h(t|X)= h0(t)exp(-0.931trt + 1.3902logWBC)

h(t|X male_trt) = h0(t)exp(-0.931 + 1.3902logWBC)

h(t|X male_con) = h0(t)exp(1.3902logWBC)

hazard ratio for male = 0.3942

female should be the same.



